### 2021-06-10 paper/paper.tex
correct incorrect results

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Dense depth estimation from image pairs}

\author{\IEEEauthorblockN{Lucas Payne}
\IEEEauthorblockA{\textit{Department of Computer Science and Software Engineering} \\
\textit{University of Canterbury}\\
Christchurch, New Zealand \\
lcp35@uclive.ac.nz}}

\maketitle

\begin{abstract}
Abstract
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
...

\section{Dense variational methods}
Varitional methods

\subsection{Dense optical flow with the Horn-Schunck method}
A well-known 1981 paper on optical flow estimation by Horn and Schunck \ref{horn_schunck} introduced what is now called the ``Horn-Schunck method'', one of the
first widespread variational algorithms used by computer vision community.
To estimate a dense field of motion vectors,

\subsection{Total-variation denoising with the Rudin-Osher-Fatemi (ROF) model}
The problem of image denoising can be put in a global optimization
framework. A cost function is formulated that penalizes ``noise'' and rewards closeness to the original (noisy) image. The algorithm then
consists of minimizing this cost function over all possible candidate ``denoised'' images.

Let $I:[0,1]^2 \rightarrow \mathbb{R}$ be a square grayscale image with continuous domain, and
$I_{i,j}$ denote the intensity at pixel $(i, j)$ in the sampled discrete image.
One formulation of the continuous ROF cost function is
\begin{equation}\label{rof_continuous}
\begin{split}
    E(\hat{I}) &= \int_0^1\int_0^1 \frac{1}{2} \left(I(x,y) - \hat{I}(x,y)\right)^2 + \lambda \|\nabla \hat{I}(x,y)\| \,dx \,dy.
\end{split}
\end{equation}
The left-hand term in the integrand is the quadratic data term, penalizing differences from the original noisy image.
$\|\nabla \hat{I}(x,y)\|$ is a measure of the local variation of intensity at a point in the image.

A common finite difference approximation for $\nabla\hat{I}(x,y)$, valid for non-boundary pixels, is
\begin{equation}
    \hat{\nabla}I_{i,j} = \left(\frac{I_{i+1,j} - I_{i-1,j}}{2\Delta x}, \frac{I_{i,j+1} - I_{i,j-1}}{2\Delta y}\right)^T,
\end{equation}
where $\Delta x, \Delta y$ are pixel extents in the image domain. Where the original image is discontinuous, such as at edges,
this finite difference vector can be very large. Furthermore, the set of pixels whose finite-difference stencils extend over discontinuities is
\textit{not} neglible in the finite approximation of integral \eqref{rof_continuous}. A quadratic regularizer, such as
$\frac{\lambda}{2} \|\nabla \hat{I}\|^2$ will harshly penalize the appearance of sharp discontinuities, since a large value returned
by a finite difference is squared. The main idea behind the ROF model is to use the non-squared ``total-variation'' $\lambda \|\nabla \hat{I}\|$.
While this is notably more difficult to optimize (precluding the use of simple linear least squares), the final effect is a preservation
of isolated discontinuities such as edges and stripe patterns, while still penalizing large patches of interior noise. Notably, the ROF model can be solved by a non-linear diffusion process, performing gradient descent to solve the Euler-Lagrange
equations of the cost functional. See their classic paper \ref{rof} for details, and \ref{cremers_rof} for a more recent discussion.

Fundamentally, dense variational methods such as \eqref{horn_schunck} and \eqref{dense_geometry} follow these same lines. First,
a cost functional of a image with continuous domain is formulated, penalizing unwanted properties of the solution. This cost functional is discretized,
and the algorithm outputs a discrete function (such as a denoised image, or a depth map, or an optical flow field) which minimizes the discrete cost function.
The majority of the complexity is in the method used to minimize (or attempt to minimize) this cost function, which could be highly non-linear.
See the TUM lecture series by Cremers \ref{variational_lectures}, freely available online, for more discussion.

% Two-dimensional areas of an image with high local variation
% are considered ``noisy''. Edges cause extremely high local variation --- in fact, an image with continuous domain may not be differentiable. However,
% finite difference approximations (such as the common ``forward difference''
% The important contribution of the ROF model
% is that this regularizer, which penalizes noise, is \textit{not} quadratic. A

\subsection{Total-variation for dense optical flow estimation}
In \ref{tv_optical_flow}, Zach et. al formulate the optical flow problem \ref{optical_flow_reference}
as a global optimization with non-quadratic data and regularizer terms. This gives a modification
of the Horn-Schunck algorithm which is more robust to outliers and which tends to preserve motion discontinuities across object boundaries.
Zach et. al's main contribution is their method of optimization,
which we will reproduce in the context of dense depth map estimation.
See their paper \ref{tv_optical_flow} for full details in the context of dense optical flow estimation.

\section{Total variation for depth map estimation from image pairs}
The paper by Cremer's et. al \ref{dense_geometry} applies the method of Zach et. al \ref{tv_optical_flow} to
a to the problem of dense depth-map
reconstruction from collections of (grayscale) images. For simplicity, we restrict our attention to the case of two images only ---
see \ref{dense_geometry} for details on the generalization. We assume (as in \ref{dense_geometry}) that we have an accurate (and constant) intrinsic camera matrix,
our images are undistorted, and, importantly, that we have accurate camera pose estimations.
Our fully non-linear cost function is
\begin{equation}
\begin{split}
E(h) &= \lambda \int_{\Omega_0}\left\|I_1(\pi(\exp(\hat{\eta}_1)X(x,h)) - I_0(\pi(x))\right\|
     + \|\nabla h\| d^2 x,
\end{split}
\end{equation}
which has TV-$L^1$ \ref{tv_optical_flow} data and regularizer terms.
(---REDO reprojection notation)
Let $\rho(h) = \left\|I_1(\pi(\exp(\hat{\eta}_1)X(x,h)) - I_0(\pi(x))\right\|$. Since the data term is non-quadratic, we
cannot apply the ROF optimization method directly. However, we can introduce
an auxiliary depth map $h^\prime$ to separate the data and regularizer terms,
and introduce a penalizer for disparity between $h^\prime$ and $h$. The
new separable cost function is
\begin{equation}\label{separable_cost}
    \hat{E}_\theta(h, \hat{h}) = \int_\Omega \lambda \|\rho(h)\| + \frac{1}{2\theta}(h - \hat{h})^2 + \|\nabla \hat{h}\|.
\end{equation}
The parameter $\theta$ is set to some small constant.
The cost function $E_\theta$ is separated into functions of $h$ and $h^\prime$,
and $h$ and $h^\prime$ are alternately optimized in the following way:
\begin{itemize}
    \item For $h$ fixed, solve
        \begin{equation}\label{minh}
            \min_{h^\prime} \int_\Omega \lambda \|\nabla h^\prime\| + \frac{1}{2\theta}(h - h^\prime)^2\,dx
        \end{equation}
    using the ROF optimization methods.
    \item For $h^\prime$ fixed, solve
        \begin{equation}\label{minhhat}
            \min_h \int_\Omega \|\rho(h)\| + \frac{1}{2\theta}(h - h^\prime)^2\,dx.
        \end{equation}
        Since $\rho(h)$ measures point-wise reprojection error, this
        optimization can be done point-wise.
\end{itemize}
The ROF cost function \eqref{minh} can be minimized using the iterative method of Chambolle \ref{chambolle}. For completeness, we reproduce this
method here in the context of dense depth map estimation. We provide no proofs --- see \ref{chambolle} for full details.

% The quadratic term $\frac{1}{2\theta}(h-h^\prime)^2$, for small $\theta$, makes
% sure that the iteration makes ``small steps''.

\section{Visualization and dataset}
A visual tool for rapidly prototyping new dense reconstruction methods is provided.
We evaluate our method on the ``freiburg3\_long\_office\_household'' RGBD dataset, which contains accurate ground-truth camera poses, and
depth maps constructed using a Kinect sensor \ref{dataset} \ref{largescale}. Our tool allows these depth maps to be used as a drop-in replacement for the estimated
depth maps, to get a rough comparison against the ``ground truth''. Figure (ref figure) shows an example visual inspection of the output
of our total-variation algorithm.

\begin{figure}[htbp]
\begin{subfigure}[b]{0.5\textwidth}
\centerline{\includegraphics[width=\textwidth]{figures/cloud_groundtruth.png}}
\caption{groundtruth}
\label{cloud_groundtruth}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\centerline{\includegraphics[width=\textwidth]{figures/cloud_computed.png}}
\caption{computed with regularizer}
\label{cloud_computed}
\end{subfigure}
\end{figure}

Figure (fig) displays some preliminary results, using heuristically determined parameters of $\theta$, $\lambda$, and $\tau$, and around 25 iterations.
Clearly this depth map is not satisfactory. However, it has some important qualitative features. Firstly, self-occlusion is captured for the
orange globe in the reprojection image, which is not captured by the brute-force method. Secondly, the point cloud clusters generally
toward the real object locations --- note the orange cluster near the globe, and the

\begin{figure}[htbp]
\begin{subfigure}[b]{0.5\textwidth}
\centerline{\includegraphics[width=\textwidth]{figures/noisevis_groundtruth.png}}
\caption{groundtruth}
\label{noisevis_groundtruth}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\centerline{\includegraphics[width=\textwidth]{figures/noisevis_computed.png}}
\caption{brute-force computed, with no regularizer}
\label{noisevis_computed}
\end{subfigure}
\end{figure}

Consider the naive approach of minimizing reprojection error pointwise. For one pixel in frame 1, the depth value parameterizes a ray which
is then reprojected into frame 2. A simple brute-force approach is to search frame 2, along this reprojected ray, for the color which gives minimum
reprojection error.
Figures \ref{noisevis_groundtruth} and \ref{noisevis_computed} visualize the problem with this approach.
Figure \ref{noisevis_groundtruth} displays, using the the ground truth depth map data provided by the Kinect sensor \ref{tum_dataset}, anticlockwise
from the bottom right:
\begin{itemize}
    \item The reprojection image.
    \item The reprojection error.
    \item A 3D view of the depth map placed in world space.
    \item The depth map.
\end{itemize}
Figure \ref{noisevis_computed} displays the same results when using a depth map computed with the naive brute-force algorithm outlined above.

The ground truth reprojection image contains ``coloured shadows'' due to occlusion. These also contribute
to high (white) values in the reprojection error image. The brute-force depth map avoids this occlusion problem.

Consider a pixel in frame 1 corresponding a point which, from the perspective of camera 2, is behind the blue die.
Likely, the line search in frame 2 will find
an unoccluded point, of similar colour, on the table. This, however, gives an unrealistic (and noisy) depth value in general --- compare
the depth images in figures \ref{noisevis_groundtruth} and \ref{noisevis_computed}. Clearly we would like to favour the depth map in
\eqref{noisevis_groundtruth}, and this is exactly the reason for the (denoising) regularizing term in \eqref{rof_continuous}.

Due to the freedom of the line search, the brute-force method has achieved low reprojection error (bottom right of figure \ref{noisevis_computed}).
With the ground truth depth map, however, the self-occlusion causes a large amount of reprojection error, which will penalize the cost function
\eqref{rof_continuous}. We would clearly like to minimize the total reprojection error (with the data term), but also keep the depth map ``smooth'' (with the regularizer term), as achieved in \ref{dense_geometry}.
However, it is difficult to choose a good value of $\lambda$ to balance the data and regularizer.

We propose a modification to the method of \ref{dense_geometry}, by removing the

\section{Comparison of results}
\section{Possibilities for further work}

\section*{References}
...

\begin{thebibliography}{00}
\bibitem{rof}
    ROF
\bibitem{cremers_rof}
    Cremers ROF
\bibitem{dense_geometry}
    Cremers, Real-Time Dense Geometry from a Handheld Camera.
\bibitem{tv_optical_flow}
    Paper in Pattern recognition book
\bibitem{dataset}
    Cremers, dataset
\bibitem{largescale}
    Cremers, largescale use of dataset
\bibitem{horn_schunck}
    Horn, Schunck.
\bibitem{volumetric}
    B. Curless, M. Levoy, A Volumetric Method for Building Complex Models from Range Images. In SIGGRAPH, 1996.
\bibitem{TUM_volumetric_CPU}
    F. Steinbr\"ucker, J. St\"urm, D. Cremers, Volumetric 3D Mapping in Real-Time on a CPU. In ICRA, 2014.
\bibitem{TUM_volumetric_octree}
    F. Steinbr\"ucker, C. Kerl, J. St\"urm, D. Cremers, Large-Scale Multi-Resolution Surface Reconstruction from RGB-D Sequences. In ICCV, 2013.
\bibitem{kinectfusion}
    R. Newcombe et al., KinectFusion: Real-Time Dense Surface Mapping and Tracking. Microsoft Research, 2011.
\bibitem{tum_dataset}
    A Benchmark for the Evaluation of RGB-D SLAM Systems (J. Sturm, N. Engelhard, F. Endres, W. Burgard and D. Cremers), In Proc. of the International Conference on Intelligent Robot Systems (IROS), 2012. 
\bibitem{variational_lectures}
    variational lecture series
\bibitem{optical_flow_reference}
    optical flow reference
\bibitem{chambolle}
    Chambolle 2004
\end{thebibliography}

\end{document}